---
title: "Assignment 5: Data transformation and visualization - Part 2"
format: gfm
editor: visual
---

```{r, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

First, load all the required packages with the following code. Install them if they are not installed yet.

```{r}
library(tidyverse)
library(knitr)
library(gapminder)
```

<br>

## Exercise 1. Trends in land value

This excercise uses a dataset that describes the trends in land value (`Land.Value`), among other variables, in different states in the US 1975-2013. The states are grouped into four different regions, under the variable `region`. This dataset was obtained from the Data Science Services of Harvard University.

```{r}
housing <- read_csv("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/landdata_states.csv")
housing %>%
  head() %>% 
  kable()
```

<br>

**1.1 Washington DC was not assigned to a region in this dataset. According to the United States Census Bureau, however, DC is part of the South region. Here:**

-   Change the region of DC to “South” (Hint: there are multiple ways to do this, but `mutate()` and `ifelse()` might be helpful)

-   Create a new tibble or regular dataframe consisting of this new updated `region` variable along with the original variables `State`, `Date` and `Land.Value` (and no others)

-   Pull out the records from DC in this new data frame. How many records are there from DC? Show the first 6 lines.

```{r}
housing |> 
  mutate(region = ifelse(is.na(region), "South", region)) |> 
  select(region, State, Land.Value, Date) |> 
  filter(State == "DC") |> 
  head() |> 
  kable()
```

Answer: 153 records. <br>

**1.2 Generate a tibble/dataframe that summarizes the mean land value of each region at each time point and show its first 6 lines.**

```{r}
housing |> 
  mutate(region = ifelse(is.na(region), "South", region)) |> 
  group_by(Date, region) |> 
  summarize(mean_land_value = mean(Land.Value)) |> 
  select(region, Date, mean_land_value) |> 
  arrange(region, Date) |> 
  head() |> 
  kable()
```

<br>

**1.3 Using the tibble/dataframe from 1.2, plot the trend in mean land value of each region through time.**

```{r}
housing |> 
  mutate(region = ifelse(is.na(region), "South", region)) |> 
  group_by(Date, region) |> 
  summarize(mean_land_value = mean(Land.Value)) |> 
  select(region, Date, mean_land_value) |> 
  arrange(region, Date) |> 
  ggplot() +
  geom_line(mapping = aes(x = Date, y = mean_land_value, color = region))
```

<br>

## Exercise 2. Life expectancy and GDP per capita 1952-2007

This exercise uses the `gapminder` dataset from the `gapminder` package. It describes the life expectancy (`lifeExp`), GDP per capita (`gdpPercap`), and population (`pop`) of 142 countries from 1952 to 2007. These countries can be grouped into 5 continents. As a reminder, **reproduce the following plots exactly as shown**.

```{r}
gapminder |> 
  head() |> 
  kable()
```

<br>

**2.1 Use a scatterplot to explore the relationship between per capita GDP (gdpPercap) and life expectancy (lifeExp) in the year 2007.**

```{r}
gapminder |> 
  filter(year == "2007") |> 
  ggplot() +
  geom_point(mapping = aes(x = gdpPercap, y = lifeExp))
```

<br>

**2.2 Add a smoothing line to the previous plot.**

```{r}
gapminder |> 
  filter(year == "2007") |> 
  ggplot() +
  geom_point(mapping = aes(x = gdpPercap, y = lifeExp)) +
  geom_smooth(mapping = aes(x = gdpPercap, y = lifeExp))
```

<br>

**2.3 Exclude Oceania from the previous plot, show each continent in a different color, and fit a separate smoothing line to each continent to identify differences in this relationship between continents. Turn off the confidence intervals.**

Note: only two Oceanian countries are included in this dataset, and `geom_smooth()` does not work with two data points, which is why they are excluded.

```{r}
gapminder |> 
  filter(year == "2007", continent != "Oceania") |> 
  ggplot() +
  geom_point(mapping = aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_smooth(mapping = aes(x = gdpPercap, y = lifeExp, color = continent), se = FALSE)
```

<br>

**2.4 Use faceting to solve the same problem. Include the confidence intervals in this plot.**

```{r}
gapminder |> 
  filter(year == "2007", continent != "Oceania") |> 
  ggplot() +
  facet_wrap(~continent) +
  geom_point(mapping = aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_smooth(mapping = aes(x = gdpPercap, y = lifeExp, color = continent))
```

<br>

**2.5 Explore the trend in life expectancy through time in each continent. Color by continent.**

```{r}
gapminder |> 
  ggplot() +
  facet_wrap(~continent) +
  geom_line(mapping = aes(x = year, y = lifeExp, color = continent, group = country))
```

<br>

**2.6 From the previous plot, we see some abnormal trends in Asia and Africa, where the the life expectancy in some countries sharply dropped at certain time periods. Here, we look into what happened in Asia in more detail. First, create a new dataset by filtering only the Asian countries. Show the first 6 lines of this filtered dataset.**

```{r}
gapminder_asia <- gapminder |> 
  filter(continent == "Asia")

gapminder_asia |> 
  head() |> 
  kable()
```

<br>

**2.7 Using the filtered dataset, identify the countries that had abnormal trends in life expectancy by plotting, and discuss historical events possibly explaining these trends. (Hint: facet by country)**

```{r}
gapminder_asia |> 
  ggplot() +
  facet_wrap(~country) +
  geom_line(mapping = aes(x = year, y = lifeExp, color = country))
```

Answer: Cambodia, China, Iraq, and North Korea (Korea, Dem. Rep.) all had abnormal dips in their life expectancy. Cambodia's trend can be explained by the civil war happening at the time. China's trend can be explained. bythe famine and political revolution happening in the 1960's. Iraq's trend can be explained by the beginning of conflict with the United States in the late 90's, early 2000's. The trend in North Korea can be explained by flooding that damaged crops and infrastructure, which caused famine.

<br>

------------------------------------------------------------------------

<br>

#### **THE LAST FOUR QUESTIONS ARE OPTIONAL**

<br>

**2.8 Explore the trend in per capita GDP through time on each continent (OPTIONAL)**

**2.9 There is one Asian country that had a very sharp decline in per capita GDP. With the previously filtered dataset, use a plot to identify this country and speculate on the historical event underlying this pattern (OPTIONAL)**

Answer: Write your response here.

**2.10 Starting from the full `gapminder` dataset, explore the relationship between per capita GDP (`gdpPercap`) and life expectancy (`lifeExp`) in 1952 and in 2007 using a scatter plot as shown below. Color the the points that represent different countries based on the continent that they are in, and make the size of the points proportional to the population size (OPTIONAL)**

**2.11 Polish the previous plot to make it more informative and nicer looking. Below is an example of what you can do, but you don’t have to completely replicate this (OPTIONAL)**
